### 2010-12-19 paper/paper.tex
Add the initial outline of the paper.

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{dependencies/acm_proc_article-sp}

\usepackage{url}
\usepackage{color}
\usepackage{verbatim}
\usepackage{listings}
\lstset{
  language=C++,             % choose the language of the code
  basicstyle=\small,       % the size of the fonts that are used for the code
  numbers=left,                   % where to put the line-numbers
  numberstyle=\footnotesize,      % the size of the fonts that are used for the line-numbers
  stepnumber=0,                   % the step between two line-numbers. If it is 1 each line will be numbered
  numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
  showspaces=false,               % show spaces adding particular underscores
  showstringspaces=false,         % underline spaces within strings
  showtabs=false,                 % show tabs within strings adding particular underscores
  %frame=single,                   % adds a frame around the code
  tabsize=2,              % sets default tabsize to 2 spaces
  captionpos=t,                   % sets the caption-position to bottom
  breaklines=false,        % sets automatic line breaking
  breakatwhitespace=false,    % sets if automatic breaks should only happen at whitespace
  escapeinside={\%}{)}          % if you want to add a comment within your code
}

\begin{document}

\title{ Development Of Parallel Matrix Operations }
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nicholas Dexter \\
       \affaddr{RIT}\\
       \email{nxd3734@rit.edu}
% 2nd. author
\alignauthor
Brian Gianforcaro \\
       \affaddr{RIT}\\
       \email{bjg1955@rit.edu}
% 3rd. author
\alignauthor
Jerome Marhic \\
       \affaddr{RIT}\\
       \email{jxm2644@rit.edu}
}
\maketitle
\begin{abstract}
%the formatting guidelines for ACM SIG Proceedings.
%It complements the document \textit{Author's Guide to Preparing
%ACM SIG Proceedings Using \LaTeX$2_\epsilon$\ and Bib\TeX}. This
%source file has been written with the intention of being
%compiled under \LaTeX$2_\epsilon$\ and BibTeX.
%
%The developers have tried to include every imaginable sort
%of ``bells and whistles", such as a subtitle, footnotes on
%title, subtitle and authors, as well as in the text, and
%every optional component (e.g. Acknowledgments, Additional
%Authors, Appendices), not to mention examples of
%equations, theorems, tables and figures.
%
%To make best use of this sample document, run it through \LaTeX\
%and BibTeX, and compare this source code with the printed
%output produced by the dvi file.
\end{abstract}
u
%% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%%%A category including the fourth, optional field follows...
%
%\terms{Theory}

\keywords{ MongoDB, NoSQL, RIT, Database System Implementation, XPath, GeoSpatial Search, Building Index's in Parallel} % NOT required for Proceedings
%
%\footnotetext[1]{NoSQL databases, also known as structured storage are known for not requiring a fixed table schema and often avoid join operations. The thing that
% links them together is there use of alternate query languages from SQL.}

\section{Introduction}
\section{}
\subsection{}
\subsection{Inserting Records Into The Database}

\subsection{Key Values And Insertion Into Indexes}
\subsection{Page Fetching}
\subsection{Query Processing And Optimizations}
\subsection{Building Indexes In Parallel}
\subsection{XPath Like Queries Of Documents}
\subsection{Expanding Geo-spatial Search}
\subsection{User Authentication}
\newpage

\newpage
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{bibliography}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
\balancecolumns
% That's all folks!
\end{document}

### 2010-12-19 src/MatrixDouble.java
Initial version of matrix classes.

@@ -1,0 +1,58 @@
+import java.util.Random;
+
+public class MatrixDouble {
+
+  private final int rows;
+  private final int cols;
+  public final double[][] data;
+
+  public MatrixDouble( int rows, int cols ) {
+    this.rows = rows;
+    this.cols = cols;
+
+    data = new double[rows][cols];
+  }
+
+  public MatrixDouble( double[][] cdata ) {
+    this.rows = cdata.length;
+    this.cols = cdata[0].length;
+
+    data = new double[rows][cols];
+    this.copyData( cdata );
+  }
+
+  private MatrixDouble( MatrixDouble copy ) { 
+    this( copy.data );
+  }
+
+  public int rows() { return rows; }
+
+  public int cols() { return cols; }
+
+  private void copyData( double[][] cdata ) {
+    for ( int r = 0; r < rows ; r++ ) {
+      for ( int c = 0; c < cols; c++ ) {
+        this.data[r][c] = cdata[r][c];
+      }
+    }
+  }
+
+  public static MatrixDouble random( int row, int col ) {
+
+    MatrixDouble rand = new MatrixDouble( row, col ); 
+
+    Random prng = new Random( 1000 );
+    
+    for ( int r = 0; r < rand.rows(); r++ ) {
+      for ( int c = 0; c < rand.cols(); c++ ) {
+        rand.data[r][c] = prng.nextDouble();
+      }
+    }
+
+    return rand;
+  }
+
+  public void display() {
+
+  }
+}

### 2010-12-19 src/MatrixInt.java

@@ -1,0 +1,60 @@
+
+import java.util.Random;
+
+public class MatrixInt {
+
+  private final int rows;
+  private final int cols;
+  public final int[][] data;
+
+  public MatrixInt( int rows, int cols ) {
+    this.rows = rows;
+    this.cols = cols;
+
+    data = new int[rows][cols];
+  }
+
+  public MatrixInt( int[][] cdata ) {
+    this.rows = cdata.length;
+    this.cols = cdata[0].length;
+
+    data = new int[rows][cols];
+    this.copyData( cdata );
+  }
+
+  private MatrixInt( MatrixInt copy ) { 
+    this( copy.data );
+  }
+
+  public int rows() { return rows; }
+
+  public int cols() { return cols; }
+
+  private void copyData( int[][] cdata ) {
+    for ( int r = 0; r < rows ; r++ ) {
+      for ( int c = 0; c < cols; c++ ) {
+        this.data[r][c] = cdata[r][c];
+      }
+    }
+  }
+
+  public static MatrixInt random( int row, int col ) {
+
+    MatrixInt rand = new MatrixInt( row, col ); 
+
+    Random prng = new Random( 1000 );
+    
+    for ( int r = 0; r < rand.rows(); r++ ) {
+      for ( int c = 0; c < rand.cols(); c++ ) {
+        rand.data[r][c] = prng.nextInt();
+      }
+    }
+
+    return rand;
+  }
+
+  public void display() {
+
+  }
+  
+}

### 2010-12-19 src/MatrixOp.java

@@ -1,0 +1,73 @@
+
+public class MatrixOp { 
+
+  // Addition Operators
+
+  public static MatrixDouble add( int value, MatrixDouble matrix ) {
+    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
+
+    for ( int r = 0; r < matrix.rows(); r++ ) {
+      for ( int c = 0; c < matrix.cols(); c++ ) {
+        output.data[r][c] = matrix.data[r][c] + value;
+      }
+    }
+    return output;
+  }
+
+  // Multiplication Operators
+
+  public static MatrixDouble mult( MatrixDouble matrix, int value ) {
+    return mult( value, matrix );
+  }
+  public static MatrixDouble mult( int value, MatrixDouble matrix  ) {
+    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
+
+    for ( int r = 0; r < matrix.rows(); r++ ) {
+      for ( int c = 0; c < matrix.cols(); c++ ) {
+        output.data[r][c] = matrix.data[r][c] * value;
+      }
+    }
+    return output;
+  }
+
+  public static MatrixDouble mult( MatrixDouble matrix, double value ) {
+    return mult( value, matrix );
+  }
+  public static MatrixDouble mult( double value, MatrixDouble matrix  ) {
+    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
+
+    for ( int r = 0; r < matrix.rows(); r++ ) {
+      for ( int c = 0; c < matrix.cols(); c++ ) {
+        output.data[r][c] = matrix.data[r][c] * value;
+      }
+    }
+    return output;
+  }
+
+  public static MatrixDouble mult( double value, MatrixInt matrix ) {
+    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
+
+    for ( int r = 0; r < matrix.rows(); r++ ) {
+      for ( int c = 0; c < matrix.cols(); c++ ) {
+        output.data[r][c] = matrix.data[r][c] * value;
+      }
+    }
+    return output;
+  }
+
+  public static MatrixInt mult( int value, MatrixInt matrix ) {
+    MatrixInt output = new MatrixInt( matrix.rows(), matrix.cols() );
+
+    for ( int r = 0; r < matrix.rows(); r++ ) {
+      for ( int c = 0; c < matrix.cols(); c++ ) {
[…]

### 2010-12-20 src/MatrixInt.java
Documentation for MatrixInt, add some tests in the main method.

@@ -1,12 +1,21 @@
 
 import java.util.Random;
 
+/**
+ * A integer based matrix
+ */
 public class MatrixInt {
 
   private final int rows;
   private final int cols;
-  public final int[][] data;
+  public int[][] data;
 
+  /**
+   * Integer Matrix constructor for an new empty object.
+   *
+   * @param rows How many rows in the matrix
+   * @param cols How many columns in the matrix
+   */
   public MatrixInt( int rows, int cols ) {
     this.rows = rows;
     this.cols = cols;
@@ -14,6 +23,11 @@ public class MatrixInt {
     data = new int[rows][cols];
   }
 
+  /**
+   * Data copy constructor
+   *
+   * @param cdata The data to copy
+   */
   public MatrixInt( int[][] cdata ) {
     this.rows = cdata.length;
     this.cols = cdata[0].length;
@@ -22,14 +36,25 @@ public class MatrixInt {
     this.copyData( cdata );
   }
 
+  /**
+   * Matrix copy constructor
+   */
   private MatrixInt( MatrixInt copy ) { 
     this( copy.data );
   }
 
+  /** Get the number of rows in the matrix */
   public int rows() { return rows; }
 
+  /** Get the number of columns of in the matrix */
   public int cols() { return cols; }
 
+  /** 
+   * Copy the data from a given array into 
+   * the matrix own allocated data.
+   *
+   * @param cdata The multidemensional array to copy from
+   */
   private void copyData( int[][] cdata ) {
     for ( int r = 0; r < rows ; r++ ) {
       for ( int c = 0; c < cols; c++ ) {
@@ -38,11 +63,16 @@ public class MatrixInt {
     }
   }
 
+  /**
+   * Generates a new matrix of the specified size, filled with random data.
+   *
+   * @param row The number of rows in the matrix
+   * @param col The number of columns in the matrix
+   * @return The new random Integer Matrix
+   */
   public static MatrixInt random( int row, int col ) {
-
     MatrixInt rand = new MatrixInt( row, col ); 
-
-    Random prng = new Random( 1000 );
+    Random prng = new Random();
     
     for ( int r = 0; r < rand.rows(); r++ ) {
       for ( int c = 0; c < rand.cols(); c++ ) {
@@ -53,8 +83,48 @@ public class MatrixInt {
     return rand;
   }
 
+  /**
+   * Generates a new matrix of the specified size, filled with random data.
+   *
[…]

### 2010-12-20 src/MatrixDouble.java
Implement MatrixDoube::display

@@ -53,6 +53,11 @@ public class MatrixDouble {
   }
 
   public void display() {
-
+    for ( int r = 0; r < rows(); r++ ) {
+      for ( int c = 0; c < cols(); c++ ) {
+        System.out.printf( "%.2f ", data[r][c] );
+      }
+      System.out.println();
+    }
   }
 }

### 2010-12-20 src/MatrixOp.java
Flesh our way more of MatrixOp, add javadocs

@@ -1,73 +1,192 @@
+/**
+ * Serial Matrix Operation Wrappers
+ */
+public class MatrixOp {
 
-public class MatrixOp { 
+  /**
+   * Add a double matrix with a scalar integer value.
+   * 
+   * @param m Matrix of type double 
+   * @param v The integer value to add the matrix with
+   * @return The resulting matrix
+   */
+  public static MatrixDouble add(MatrixDouble m, int v) {
+    return add(v,m);
+  }
 
-  // Addition Operators
-
-  public static MatrixDouble add( int value, MatrixDouble matrix ) {
-    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
-
-    for ( int r = 0; r < matrix.rows(); r++ ) {
-      for ( int c = 0; c < matrix.cols(); c++ ) {
-        output.data[r][c] = matrix.data[r][c] + value;
+  /**
+   * Add a double matrix with a scalar integer value.
+   * 
+   * @param v The integer value to add the matrix by
+   * @param m Matrix of type double 
+   * @return The resulting matrix
+   */
+  public static MatrixDouble add(int v, MatrixDouble m) {
+    MatrixDouble output = new MatrixDouble(m.rows(), m.cols());
+    for (int r = 0; r < m.rows(); r++) {
+      for (int c = 0; c < m.cols(); c++) {
+        output.data[r][c] = m.data[r][c] + v;
       }
     }
     return output;
   }
 
-  // Multiplication Operators
-
-  public static MatrixDouble mult( MatrixDouble matrix, int value ) {
-    return mult( value, matrix );
-  }
-  public static MatrixDouble mult( int value, MatrixDouble matrix  ) {
-    MatrixDouble output = new MatrixDouble( matrix.rows(), matrix.cols() );
-
-    for ( int r = 0; r < matrix.rows(); r++ ) {
-      for ( int c = 0; c < matrix.cols(); c++ ) {
-        output.data[r][c] = matrix.data[r][c] * value;
-      }
-    }
-    return output;
+  /**
+   * Add a integer matrix with a scalar integer value.
+   * 
+   * @param m Matrix of type integer
+   * @param v The integer value to add the matrix with
+   * @return The resulting matrix
+   */
+  public static MatrixInt add(MatrixInt m, int v) {
+    return add(v,m);
   }
 
-  public static MatrixDouble mult( MatrixDouble matrix, double value ) {
-    return mult( value, matrix );
-  }
[…]

### 2010-12-20 src/MatrixDouble.java
Initial parallel matrix multiplication

@@ -6,6 +6,13 @@ public class MatrixDouble {
   private final int cols;
   public final double[][] data;
 
+  /**
+   * Double Matrix constructor for an new empty object.
+   *
+   * @param rows How many rows in the matrix
+   * @param cols How many columns in the matrix
+   */
+
   public MatrixDouble( int rows, int cols ) {
     this.rows = rows;
     this.cols = cols;
@@ -13,6 +20,11 @@ public class MatrixDouble {
     data = new double[rows][cols];
   }
 
+  /**
+   * Data copy constructor
+   *
+   * @param cdata The data to copy
+   */
   public MatrixDouble( double[][] cdata ) {
     this.rows = cdata.length;
     this.cols = cdata[0].length;
@@ -21,14 +33,25 @@ public class MatrixDouble {
     this.copyData( cdata );
   }
 
+  /**
+   * Matrix copy constructor
+   */
   private MatrixDouble( MatrixDouble copy ) { 
     this( copy.data );
   }
 
+  /** Get the number of rows in the matrix */
   public int rows() { return rows; }
 
+  /** Get the number of columns of in the matrix */
   public int cols() { return cols; }
 
+  /** 
+   * Copy the data from a given array into 
+   * the matrix own allocated data.
+   *
+   * @param cdata The multidemensional array to copy from
+   */
   private void copyData( double[][] cdata ) {
     for ( int r = 0; r < rows ; r++ ) {
       for ( int c = 0; c < cols; c++ ) {
@@ -37,6 +60,13 @@ public class MatrixDouble {
     }
   }
 
+  /**
+   * Generates a new matrix of the specified size, filled with random data.
+   *
+   * @param row The number of rows in the matrix
+   * @param col The number of columns in the matrix
+   * @return The new random Integer Matrix
+   */
   public static MatrixDouble random( int row, int col ) {
 
     MatrixDouble rand = new MatrixDouble( row, col ); 
@@ -52,6 +82,36 @@ public class MatrixDouble {
     return rand;
   }
 
+  /**
+   * Generates a new matrix of the specified size, filled with random data.
+   *
+   * @param row The number of rows in the matrix
+   * @param col The number of columns in the matrix
+   * @return The new random Integer Matrix
+   */
+  public static MatrixDouble randomSmall( int row, int col ) {
+    double REDUCTION = 10000000;
[…]

### 2010-12-20 src/MatrixInt.java

@@ -16,7 +16,7 @@ public class MatrixInt {
    * @param rows How many rows in the matrix
    * @param cols How many columns in the matrix
    */
-  public MatrixInt( int rows, int cols ) {
+  public MatrixInt( final int rows, final int cols ) {
     this.rows = rows;
     this.cols = cols;
 
@@ -39,7 +39,7 @@ public class MatrixInt {
   /**
    * Matrix copy constructor
    */
-  private MatrixInt( MatrixInt copy ) { 
+  private MatrixInt( final MatrixInt copy ) { 
     this( copy.data );
   }
 
@@ -70,7 +70,7 @@ public class MatrixInt {
    * @param col The number of columns in the matrix
    * @return The new random Integer Matrix
    */
-  public static MatrixInt random( int row, int col ) {
+  public static MatrixInt random( final int row, final int col ) {
     MatrixInt rand = new MatrixInt( row, col ); 
     Random prng = new Random();
     
@@ -90,7 +90,7 @@ public class MatrixInt {
    * @param col The number of columns in the matrix
    * @return The new random Integer Matrix
    */
-  public static MatrixInt randomSmall( int row, int col ) {
+  public static MatrixInt randomSmall( final int row, final int col ) {
     int REDUCTION = 10000000;
     MatrixInt rand = MatrixInt.random( row, col );
     for ( int r = 0; r < rand.rows(); r++ ) {
@@ -99,6 +99,14 @@ public class MatrixInt {
       }
     }
     return rand;
+  }
+
+  /** Determine if two matricies are dimensionally equal */
+  public boolean equalDim( final MatrixInt x ) {
+    return x.rows() == rows() && x.cols() == cols();
+  }
+  public boolean equalDim( final MatrixDouble x ) {
+    return x.rows() == rows() && x.cols() == cols();
   }
 
 

### 2010-12-20 src/MatrixOp.java

@@ -172,15 +172,42 @@ public class MatrixOp {
     return output;
   }
 
+  /**
+   * Multiply a integer matrix by a integer matrix.
+   * 
+   * @param a The integer value to multiply the matrix by
+   * @param a The other integer value to multiply the matrix by
+   * @return The resulting matrix
+   */
+  public static MatrixInt mult( MatrixInt a, MatrixInt b ) {
+
+    if ( ! a.equalDim( b ) ) {
+      throw new RuntimeException("Multiplication of different sized matricies");
+    }
+
+    MatrixInt output = new MatrixInt(a.rows(), a.cols());
+
+    for (int r = 0; r < output.rows(); r++) {
+      for (int c = 0; c < output.cols(); c++) {
+        for (int k = 0; k < a.cols(); k++) {
+          output.data[r][c] += (a.data[r][k] * b.data[k][c]);
+        }
+      }
+    }
+    return output;
+  }
+
+
   public static void main( String args[] ) {
 
-    MatrixInt A = new MatrixInt( 10000, 10000 );
+    MatrixInt A = MatrixInt.random( 900, 900 );
+    MatrixInt B = MatrixInt.random( 900, 900 );
 
     int loop = 20;
     long start = System.currentTimeMillis();
     for (int i = 0; i < loop; i++ ) {
       long lstart = System.currentTimeMillis();
-      A = MatrixOp.mult( A, 2 );
+      MatrixInt C = MatrixOp.mult( A, B );
       System.out.println( "Running time = " +
                         (System.currentTimeMillis()-lstart) + " msec" );
     }

### 2010-12-20 src/MatrixParallelOp.java

@@ -1,0 +1,108 @@
+
+import edu.rit.pj.Comm;
+import edu.rit.pj.ParallelRegion;
+import edu.rit.pj.ParallelTeam;
+import edu.rit.pj.IntegerForLoop;
+
+
+
+/**
+ * Serial Matrix Operation Wrappers
+ */
+public class MatrixParallelOp {
+
+  /**
+   * Multiply a integer matrix by a integer matrix.
+   * 
+   * @param a The integer value to multiply the matrix by
+   * @param a The other integer value to multiply the matrix by
+   * @return The resulting matrix
+   */
+  public static MatrixInt mult( final MatrixInt a, final MatrixInt b ) throws Exception {
+
+    if ( ! a.equalDim( b ) ) {
+      throw new RuntimeException("Multiplication of different sized matricies");
+    }
+
+    final MatrixInt output = new MatrixInt(a.rows(), a.cols());
+
+    new ParallelTeam().execute( new ParallelRegion() {
+      public void run() throws Exception {
+        execute( 0, output.rows()-1, new IntegerForLoop() {
+          public void run( int first, int last ) {
+            for ( int r = first; r <= last; r++ ) {
+              for (int c = 0; c < output.cols(); c++) {
+                for (int k = 0; k < a.cols(); k++) {
+                  output.data[r][c] += (a.data[r][k] * b.data[k][c]);
+                }
+              }
+            }
+          }
+        });
+      }
+    });
+
+    return output;
+  }
+
+  public static MatrixDouble mult( final MatrixDouble b, final MatrixInt a ) throws Exception {
+    return mult( a, b );
+  }
+
+  /**
+   * Multiply a integer matrix by a double matrix.
+   * 
+   * @param a The integer value to multiply the matrix by
+   * @param a The double matrix to multiply the matrix by
+   * @return The resulting matrix
+   */
+  public static MatrixDouble mult( final MatrixInt a, final MatrixDouble b ) throws Exception {
+
+    if ( ! a.equalDim( b ) ) {
+      throw new RuntimeException("Multiplication of different sized matricies");
+    }
+
+    final MatrixDouble output = new MatrixDouble(a.rows(), a.cols());
+
+    new ParallelTeam().execute( new ParallelRegion() {
+      public void run() throws Exception {
+        execute( 0, output.rows()-1, new IntegerForLoop() {
+          public void run( int first, int last ) {
+            for ( int r = first; r <= last; r++ ) {
[…]

### 2011-02-13 src/MatrixInverse.java
Modified the solve method to be in line with the parallel version.

@@ -16,22 +16,59 @@ public class MatrixInverse{
      * @return  x    the solution
      */
     public static VectorDouble solve(MatrixDouble A, VectorDouble b) throws Exception {
-        // Row echelon form of A
-        MatrixDouble REA = eliminate(A);
+        // Determine the number of rows of A
+        int N = A.rows();
+
+        // Iterate through all of the rows
+        for(int pivot = 0; pivot < N; pivot++) {
+
+            // Initialize the maximum to the first row out of the remaining 
+            // N-pivot rows
+            int max = pivot;
+            // Iterate through the remaining rows
+            for(int i = pivot + 1; i < N; i++) {
+                // If the absolute value of this row is greater than the last max
+                if(Math.abs(A.data[i][pivot]) > Math.abs(A.data[max][pivot])) {
+                    // update the new maximum
+                    max = i;
+                }
+            }
+            // Holder for the pivot row
+            double[] holder = A.data[pivot]; 
+            // Swap the old max row with the pivot row
+            A.data[pivot] = A.data[max]; 
+            // Swap the pivot row into the old max row
+            A.data[max] = holder;
+
+            // If the leading coefficient is close to 0
+            if (Math.abs(A.data[pivot][pivot]) <= eps) {
+                // Exception because matrix may be singular
+                throw new RuntimeException("Matrix incapable of inverse" +
+                       " solution via Gaussian Elimination");
+            }
+
+            // Perform elementary row operations
+            for (int i = pivot + 1; i < N; i++) {
+                double alpha = A.data[i][pivot] / A.data[pivot][pivot];
+                for (int j = pivot; j < N; j++) {
+                    A.data[i][j] = A.data[i][j] - alpha * A.data[pivot][j];
+                }
+            }
+        }
+
         // solution vector
         VectorDouble x = new VectorDouble(b.length());
         // set all x_i to 0
         x.zero();
 
-        int N = b.length();
         // perform backsubstitution
         for (int i = N - 1; i >= 0; i--) {
             // initialize sum to 0
             double sum = 0.0;
             for (int j = i + 1; j < N; j++) {
-                sum += REA.data[i][j] * x.data[j];
+                sum += A.data[i][j] * x.data[j];
             }
-            x.data[i] = (b.data[i] - sum) / REA.data[i][i];
+            x.data[i] = (b.data[i] - sum) / A.data[i][i];
         }
         return x;
     }
@@ -71,12 +108,10 @@ public class MatrixInverse{
[…]

### 2011-02-13 src/MatrixInverseSmp.java
Made parallel and modified the solve method to be in line with sequential.

@@ -1,13 +1,21 @@
+import edu.rit.pj.Comm;
+import edu.rit.pj.ParallelRegion;
+import edu.rit.pj.ParallelTeam;
+import edu.rit.pj.IntegerForLoop;
+
 /**
  * A program that applies Gaussian eliminati n t  compute
  * the inverse that solves the system Ax=b
  */
 public class MatrixInverseSmp{
 
+    static ParallelTeam team = new ParallelTeam();
     private static final double eps = 1e-10;
 
     // prevent construction
     private MatrixInverseSmp() { }
+
+
 
     /**
      * Solves the linear system A.x=b
@@ -15,33 +23,9 @@ public class MatrixInverseSmp{
      * @param   b    the RHS
      * @return  x    the solution
      */
-    public static VectorDouble solve(MatrixDouble A, VectorDouble b) throws Exception {
-        // Row echelon form of A
-        MatrixDouble REA = eliminate(A);
-        // solution vector
-        VectorDouble x = new VectorDouble(b.length());
-        // set all x_i to 0
-        x.zero();
-
-        int N = b.length();
-        // perform backsubstitution
-        for (int i = N - 1; i >= 0; i--) {
-            // initialize sum to 0
-            double sum = 0.0;
-            for (int j = i + 1; j < N; j++) {
-                sum += REA.data[i][j] * x.data[j];
-            }
-            x.data[i] = (b.data[i] - sum) / REA.data[i][i];
-        }
-        return x;
-    }
-
-    /** 
-     * apply gaussian elimination to the matrix a
-     */
-    public static MatrixDouble eliminate(MatrixDouble A) throws Exception {
+    public static VectorDouble solve(final MatrixDouble A, final VectorDouble b) throws Exception {
         // Determine the number of rows of A
-        int N = A.rows();
+        final int N = A.rows();
 
         // Iterate through all of the rows
         for(int pivot = 0; pivot < N; pivot++) {
@@ -71,15 +55,98 @@ public class MatrixInverseSmp{
                        " solution via Gaussian Elimination");
             }
 
-            // Set leading coefficient to 1 and perform elementary row op
-            for (int i = pivot + 1; i < N; i++) {
-                // Determine the factor to reduce leading coeff to 1
-                double alpha = A.data[i][pivot] / A.data[pivot][pivot];
-                for (int j = pivot; j < N; j++) {
-                    // Reduce the row to match the operation
-                    A.data[i][j] = A.data[i][j] - alpha * A.data[pivot][j];
+
[…]

### 2011-02-14 src/MatrixInverse.java
Added missing Comm.init statement to the main method.

@@ -1,3 +1,5 @@
+import edu.rit.pj.Comm;
+
 /**
  * A program that applies Gaussian eliminati n t  compute
  * the inverse that solves the system Ax=b
@@ -124,6 +126,9 @@ public class MatrixInverse{
      * main method for testing the inversion
      */
     public static void main(String args[]) throws Exception {
+
+        Comm.init(args);
+
         /*
         double[][] data = new double[3][3];
         data[0][0] = 2; data[0][1] = 1; data[0][2] = 1;

### 2011-02-14 src/MatrixInverseSmp.java
Added missing Comm.init statement to the main method.

@@ -156,6 +156,9 @@ public class MatrixInverseSmp{
      * main method for testing the inversion
      */
     public static void main(String args[]) throws Exception {
+
+        Comm.init(args);
+
         /*
         double[][] data = new double[3][3];
         data[0][0] = 2; data[0][1] = 1; data[0][2] = 1;

### 2011-02-14 src/Tester.java
Add tester

@@ -1,0 +1,85 @@
+import edu.rit.pj.Comm;
+
+public class Tester {
+
+
+    private void runTest( int N ) {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        MatrixInt B = MatrixInt.random(N,N);
+
+        long start = System.currentTimeMillis();
+
+        MatrixInt C = MatrixOp.mult( A, B );
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Classic = ("+N+") = " +(end-start));
+    }
+
+    private void runStraTest( int N ) {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        MatrixInt B = MatrixInt.random(N,N);
+
+        long start = System.currentTimeMillis();
+
+        MatrixInt C = MatrixOp.strassenMult( A, B );
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Strassen ("+N+") = "+(end-start));
+    }
+
+    private void runPTest( int N ) {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        MatrixInt B = MatrixInt.random(N,N);
+
+        long start = System.currentTimeMillis();
+
+        MatrixInt C = MatrixParallelOp.mult( A, B );
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Parallel ("+N+") = "+(end-start));
+    }
+
+    private void runPStraTest( int N ) {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        MatrixInt B = MatrixInt.random(N,N);
+
+        long start = System.currentTimeMillis();
+
+        MatrixInt C = MatrixParallelStr.mult( A, B );
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Parallel Strassen ("+N+") = "+(end-start));
+    }
+
+
+
+
+    public static void main( String[] args ) {
+
+        Comm.init( args );
+        runTest( 100 );
+        runTest( 500 );
+        runTest( 1000 );
+        runTest( 1024 );
+
+        runStraTest( 100 );
+        runStraTest( 500 );
+        runStraTest( 1000 );
+        runStraTest( 1024 );
+
+        runPTest( 100 );
+        runPTest( 500 );
+        runPTest( 1000 );
+        runPTest( 1024 );
+
+        runPStraTest( 100 );
+        runPStraTest( 500 );
+        runPStraTest( 1000 );
+        runPStraTest( 1024 );
+    }
+
+}

### 2011-02-14 src/MatrixOp.java
bug on strassen seq and parallel Please enter the commit message for your changes.

@@ -278,8 +278,9 @@ public class MatrixOp {
    * @param nbrows
    * @param nbcols
    */
-  public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) {
-
+  public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
+  {
+	System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
     int[][] data = new int[nbrows][nbcols];
     for(int i = 0; i < a.rows(); i++ ) {
       for(int j = 0; j < a.cols(); j++ ) {
@@ -298,32 +299,55 @@ public class MatrixOp {
 
   public static MatrixInt strassenMult( MatrixInt a, MatrixInt b )
   {
+	  if ( ! a.correctDim( b ) ) {
+	      throw new RuntimeException("Multiplication of different sized matricies");
+	  }
+	  
     // check if the size is a power of 2 :
-    int rows = Math.min(a.rows(),b.rows());
-    int cols = Math.min(a.cols(),b.cols());
-    int rowsnextpow2 = rows;
-    int colsnextpos2 = cols;
+	int finalrows = a.rows();
+	int finalcol = b.cols();
+	  
+    int rows = Math.max(a.rows(),b.rows());
+    int cols = Math.max(a.cols(),b.cols());
+    int resizing = Math.max(rows,cols);
+    // we want a and b to be 2 matrices of size the closest power of 2 of 'resizing'
+	
+    MatrixInt result;
 
-    if ( ! isPow2(rows) ) {
-      rowsnextpow2 = (int)(Math.log((double) rows)/Math.log(2.0f)) + 1;
-      System.out.println(rows+" closest power = "+rowsnextpow2+" = "+Math.pow(2, rowsnextpow2) );
+    if ( a.equalDim(b)  && isPow2(rows) ) 
+    {
+        System.out.println("no need to pad");
+    	result =  strassenMultRun(a, b);
+
     }
-    if ( ! isPow2(cols) ) {
-      colsnextpos2 = (int)(Math.log((double) cols)/Math.log(2.0f)) + 1;
-      System.out.println(cols+"closest power = "+colsnextpos2+" = "+Math.pow(2, colsnextpos2) );
+    else
+    {
+    	if( ! isPow2(resizing) )
+            resizing = (int)(Math.log((double) resizing)/Math.log(2.0f)) + 1;
+
[…]

### 2011-02-14 src/MatrixParallelStr.java

@@ -14,28 +14,78 @@ public class MatrixParallelStr {
   // Create a static team to avoid allocation hit
   static ParallelTeam team = new ParallelTeam();
 
-  public static MatrixInt mult( MatrixInt a, MatrixInt b ) throws Exception {
+  
+  /** pads a matrix with 0 until it reaches the new size nbrows, nbcols
+   * 
+   * @param a
+   * @param nbrows
+   * @param nbcols
+   */
+  public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
+  {
+	System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
+    int[][] data = new int[nbrows][nbcols];
+    for(int i = 0; i < a.rows(); i++ ) {
+      for(int j = 0; j < a.cols(); j++ ) {
+        data[i][j] = a.data[i][j];
+      }
+    }
+    return new MatrixInt( data );
+  }
+  /* return true if number is a power of 2
+   */
+  public static boolean isPow2(int number)
+  {
+    return ((number ^ (number-1)) == (number+(number-1))) && (number != 0);
+  }
+  
+  public static MatrixInt mult( MatrixInt a, MatrixInt b ) throws Exception 
+  {
+	  
+	  if ( ! a.correctDim( b ) ) {
+	      throw new RuntimeException("Multiplication of different sized matricies");
+	  }
+	  
+    // check if the size is a power of 2 :
+	int finalrows = a.rows();
+	int finalcol = b.cols();
+	  
+    int rows = Math.max(a.rows(),b.rows());
+    int cols = Math.max(a.cols(),b.cols());
+    int resizing = Math.max(rows,cols);
+    // we want a and b to be 2 matrices of size the closest power of 2 of 'resizing'
+	
+    MatrixInt result;
 
-    // Check if the size is a power of 2 :
-    int rows = Math.min(a.rows(),b.rows());
-    int cols = Math.min(a.cols(),b.cols());
-    int rowsnextpow2 = rows;
-    int colsnextpos2 = cols;
+    if ( a.equalDim(b)  && isPow2(rows) ) 
+    {
+        System.out.println("no need to pad");
+    	result =  multRunSmp( a, b );
 
-    if ( !MatrixOp.isPow2( rows ) ) {
[…]

### 2011-02-14 src/Tester.java

@@ -29,7 +29,7 @@ public class Tester {
         System.out.println( "Strassen ("+N+") = "+(end-start));
     }
 
-    private void runPTest( int N ) {
+    private void runPTest( int N ) throws Exception {
 
         MatrixInt A = MatrixInt.random(N,N);
         MatrixInt B = MatrixInt.random(N,N);
@@ -42,7 +42,7 @@ public class Tester {
         System.out.println( "Parallel ("+N+") = "+(end-start));
     }
 
-    private void runPStraTest( int N ) {
+    private void runPStraTest( int N ) throws Exception {
 
         MatrixInt A = MatrixInt.random(N,N);
         MatrixInt B = MatrixInt.random(N,N);

### 2011-02-14 src/MatrixOp.java
Bring in data and tester changes

@@ -280,7 +280,7 @@ public class MatrixOp {
    */
   public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
   {
-	System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
+	//System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
     int[][] data = new int[nbrows][nbcols];
     for(int i = 0; i < a.rows(); i++ ) {
       for(int j = 0; j < a.cols(); j++ ) {
@@ -316,7 +316,7 @@ public class MatrixOp {
 
     if ( a.equalDim(b)  && isPow2(rows) ) 
     {
-        System.out.println("no need to pad");
+        //System.out.println("no need to pad");
     	result =  strassenMultRun(a, b);
 
     }

### 2011-02-14 src/MatrixParallelStr.java

@@ -23,7 +23,7 @@ public class MatrixParallelStr {
    */
   public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
   {
-	System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
+	//System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
     int[][] data = new int[nbrows][nbcols];
     for(int i = 0; i < a.rows(); i++ ) {
       for(int j = 0; j < a.cols(); j++ ) {
@@ -59,7 +59,7 @@ public class MatrixParallelStr {
 
     if ( a.equalDim(b)  && isPow2(rows) ) 
     {
-        System.out.println("no need to pad");
+        //System.out.println("no need to pad");
     	result =  multRunSmp( a, b );
 
     }
@@ -159,7 +159,7 @@ public class MatrixParallelStr {
 
     int n = a.rows();
     if ( !( a.equalDim(b)) || ! (a.cols()==n ) || ! MatrixOp.isPow2(n) ) {
-      System.out.println("matrices dimensions aren't suitable for the algorithm");
+      //System.out.println("matrices dimensions aren't suitable for the algorithm");
     }
 
     if ( n <= 128 ) { // 128 seems to be the n where classic mult is more efficient (64 is still fine)

### 2011-02-14 src/Tester.java

@@ -2,84 +2,89 @@ import edu.rit.pj.Comm;
 
 public class Tester {
 
+	  static int NUM_RUNS = 4;
 
-    private void runTest( int N ) {
-
-        MatrixInt A = MatrixInt.random(N,N);
-        MatrixInt B = MatrixInt.random(N,N);
-
-        long start = System.currentTimeMillis();
-
-        MatrixInt C = MatrixOp.mult( A, B );
-
-        long end = System.currentTimeMillis();
-        System.out.println( "Classic = ("+N+") = " +(end-start));
-    }
-
-    private void runStraTest( int N ) {
+    private static void runTest( int N ) throws Exception {
 
         MatrixInt A = MatrixInt.random(N,N);
         MatrixInt B = MatrixInt.random(N,N);
 
         long start = System.currentTimeMillis();
 
-        MatrixInt C = MatrixOp.strassenMult( A, B );
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixOp.mult( A, B );
+				}
 
         long end = System.currentTimeMillis();
-        System.out.println( "Strassen ("+N+") = "+(end-start));
+        System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
     }
 
-    private void runPTest( int N ) throws Exception {
+    private static void runStraTest( int N ) throws Exception {
 
         MatrixInt A = MatrixInt.random(N,N);
         MatrixInt B = MatrixInt.random(N,N);
 
         long start = System.currentTimeMillis();
 
-        MatrixInt C = MatrixParallelOp.mult( A, B );
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        	MatrixInt C = MatrixOp.strassenMult( A, B );
+				}
 
         long end = System.currentTimeMillis();
-        System.out.println( "Parallel ("+N+") = "+(end-start));
+        System.out.println( "Strassen ("+N+") = "+(end-start)/NUM_RUNS);
     }
 
-    private void runPStraTest( int N ) throws Exception {
+    private static void runPTest( int N ) throws Exception {
 
         MatrixInt A = MatrixInt.random(N,N);
         MatrixInt B = MatrixInt.random(N,N);
 
         long start = System.currentTimeMillis();
 
-        MatrixInt C = MatrixParallelStr.mult( A, B );
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
[…]

### 2011-02-14 src/Tester.java
data tables and tester changes

@@ -74,17 +74,17 @@ public class Tester {
         //runTest( 500 );
         //runTest( 1000 );
 
-        //runStraTest( 100 );
-        //runStraTest( 500 );
-        //runStraTest( 1000 );
+        runStraTest( 100 );
+        runStraTest( 500 );
+        runStraTest( 1000 );
 
-        runPTest( 100 );
-        runPTest( 500 );
-        runPTest( 1000 );
+        //runPTest( 100 );
+        //runPTest( 500 );
+        //runPTest( 1000 );
 
-        runPStraTest( 100 );
-        runPStraTest( 500 );
-        runPStraTest( 1000 );
+        //runPStraTest( 100 );
+        //runPStraTest( 500 );
+        //runPStraTest( 1000 );
     }
 
 }

### 2011-02-14 src/VectorInt.java

@@ -1,3 +1,4 @@
+import java.util.Random;
 
 public class VectorInt {
 
@@ -18,4 +19,21 @@ public class VectorInt {
             this.data[i] = 0;
         }
     }
+
+    /**
+     * Generates a new vector of the specified size, filled with random data.
+     *
+     * @param row The number of rows in the vector
+     * @return The new random Double Vector
+     */
+    public static VectorInt random( int row ) {
+        VectorInt rand = new VectorInt( row ); 
+        Random prng = new Random( 1000 );
+        for ( int r = 0; r < rand.length(); r++ ) {
+            rand.data[r] = prng.nextInt();
+        }
+        return rand;
+    }
+
+
 }

### 2011-02-14 src/TesterInverse.java
Add other testers

@@ -1,0 +1,49 @@
+import edu.rit.pj.Comm;
+
+public class TesterInverse {
+
+	  static int NUM_RUNS = 4;
+
+    private static void runTest( int N ) throws Exception {
+
+        MatrixDouble A = MatrixDouble.random(N,N);
+        VectorDouble B = VectorDouble.random(N);
+
+
+        long start = System.currentTimeMillis();
+
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+          VectorDouble C = MatrixInverse.solve( A, B );
+				}
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
+    }
+
+    private static void runPTest( int N ) throws Exception {
+
+        MatrixDouble A = MatrixDouble.random(N,N);
+        VectorDouble B = VectorDouble.random(N);
+
+        long start = System.currentTimeMillis();
+
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        	VectorDouble C = MatrixInverseSmp.solve( A, B );
+				}
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Parallel ("+N+") = "+(end-start)/NUM_RUNS);
+    }
+
+    public static void main( String[] args ) throws Exception {
+
+        Comm.init( args );
+        runTest( 100 );
+        runTest( 500 );
+        runTest( 1000 );
+
+        runPTest( 100 );
+        runPTest( 500 );
+        runPTest( 1000 );
+    }
+}

### 2011-02-14 src/TesterVector.java

@@ -1,0 +1,48 @@
+import edu.rit.pj.Comm;
+
+public class TesterVector {
+
+	  static int NUM_RUNS = 4;
+
+    private static void runTest( int N ) throws Exception {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        VectorInt B = VectorInt.random(N);
+
+        long start = System.currentTimeMillis();
+
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixOp.mult( A, B );
+				}
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
+    }
+
+    private static void runPTest( int N ) throws Exception {
+
+        MatrixInt A = MatrixInt.random(N,N);
+        VectorInt B = VectorInt.random(N);
+
+        long start = System.currentTimeMillis();
+
+				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        	MatrixInt C = MatrixParallelOp.mult( A, B );
+				}
+
+        long end = System.currentTimeMillis();
+        System.out.println( "Parallel ("+N+") = "+(end-start)/NUM_RUNS);
+    }
+
+    public static void main( String[] args ) throws Exception {
+
+        Comm.init( args );
+        runTest( 1000 );
+        runTest( 2000 );
+        runTest( 4000 );
+
+        runPTest( 1000 );
+        runPTest( 2000 );
+        runPTest( 4000 );
+    }
+}

### 2011-02-14 paper/paper.tex
First revision.

@@
 %output produced by the dvi file.
+
+A Parallel Matrix Manipulation Library (PMML) is presented for use on sequential and symmetric multiprocessing (SMP) systems.
+PMML is written using the Parallel Java library developed by Professor Alan Kaminsky at the Rochester Institute of Technology.
+PMML provides methods for the following operations: matrix-vector multiplication, matrix-matrix multiplication, and matrix inversion and solution to the linear system $A x=b$.
+An implementation of the Strassen algorithm for efficient matrix multiplication is provided for use sequentially and on SMP systems.
+Gaussian elimination with partial-pivoting is used to find the row-reduced form of the matrix $A$ and backsubstitution is used to solve the linear system.
+Algorithms for sequential solution and parallel solution on SMP systems are provided.
+Performance metrics are presented to compare the {\em speedup} and {\em sizeup} gained in the parallel over the sequential versions.
 \end{abstract}
-u
+
 %% A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
+%\category{H.4}{Information Systems Applications}{Miscellaneous}
 %%%A category including the fourth, optional field follows...
@@
 
-\keywords{ MongoDB, NoSQL, RIT, Database System Implementation, XPath, GeoSpatial Search, Building Index's in Parallel} % NOT required for Proceedings
+\keywords{ Parallel Matrix Manipulation Library, PMML, Parallel Java, PJ, Strassen algorithm, Gaussian elimination, RIT } % NOT required for Proceedings
 %
@@
 \section{Introduction}
-\section{}
-\subsection{}
-\subsection{Inserting Records Into The Database}
+\subsection{Numerical Linear Algebra}
 
-\subsection{Key Values And Insertion Into Indexes}
-\subsection{Page Fetching}
-\subsection{Query Processing And Optimizations}
-\subsection{Building Indexes In Parallel}
-\subsection{XPath Like Queries Of Documents}
-\subsection{Expanding Geo-spatial Search}
-\subsection{User Authentication}
+The use of computers to manipulate and solve matrices has found many applications in computer science, finances, engineering, biology, and physics.
+Since many problems can be reduced to a system of linear equations, particularly in the solution of differential equations, numerical linear algebra has become increasingly useful in the study of such systems.
+In the High Performance Computing world, a large amount of effort has been devoted to developing toolsets that allow these problems to be solved in parallel.
+
+Since its development in 1979, the {\em de facto} standard for basic linear algebra operations has been the Basic Linear Algebra Subprograms (BLAS) \cite{BLAS}.
+Most modern implementations of numerical linear algebra libraries still rely on this toolset written in FORTRAN90 \cite{LAPACK}, though other language implementations exist \cite{CLAPACK,JBLAS}.
+Early in the applications of computers to such problems, vector computers reigned supreme in numerical linear algebra due to their ability to apply operations to one-dimensional arrays of data.
+As scalar processors became cheaper and more readily available, a trend began of using large systems of scalar processors to perform parallel operations.
+
+\subsection{Parallel Matrix Operations}
+
+Mention Parallel Java Library and need for Matrix Manipulation methods.
+
+\section{Methods}
+\subsection{Matrix-Vector Multiplication}
+
+Describe the problem of matrix-vector multiplication and introduce the terms.
+
+\subsection{Matrix-Matrix Multiplication}
+
+\subsubsection{Classic matrix multiplication}
+The following code:
+\begin{verbatim}
+Matrix C = MatrixOp.mult(A,B)
+Matrix C = MatrixParallelOp.mult(A,B)
+\end{verbatim}
+returns the product $C=AB$, using respectively the sequential and parallel algorithms.
+
+Note that the operation is valid only if the number of columns of $A$ equals the number of rows of $B$, otherwise an Exception is thrown.
+
+The matrix multiplication is performed by multiplying each row of $A$ by each column of $B$ \cite{MatrixMultiplication}.
+Sequentially, one thread performs all the operations by looping on every rows of $A$ and every columns of $B$.
+The complexity is in $O(m\cdot p\cdot n)$ if $A$ and $B$ are rectangular matrices of size $m\cdot p$ and $p\cdot n$ respectively.
+This method is easily parallelizable over several threads; each thread computing a range of rows of $A$.
+The complexity is then $O( \frac{m}{K}\cdot p\cdot n )$, $K$ being the number of threads used.
+
+\subsubsection{Strassen matrix multiplication}
+The following code:
+\begin{verbatim}
+Matrix C = MatrixOp.strassenMult(A,B)
+Matrix C = MatrixParallelStr.mult(A,B)
+\end{verbatim}
+returns the product $C=AB$, using respectively the sequential and parallel versions of the Strassen algorithms.
+
+There is a way of multiplying two $2\times2$ matrices using only 7 multiplications instead of 8 with the classic algorithm.
+The Strassen algorithm takes advantage of that to multiply two matrices with a complexity of $O(n^2.807)$, providing that those matrices have the same size and are square matrices with side of size a power of 2 \cite{StrassenAlgorithm}.
+The idea consists in recursively dividing both matrices into 4 smaller square matrices, until those submatrices degenerate into numbers, that is have a size of 1, and replace the last multiplication by the sum of 18 submatrices.
+
+However, by padding with zeroes the matrices to the closest power of 2, we can multiply matrices of any size.
+
+Due to this recursive nature, this algorithm is not easily parallelizable but we managed to parallelize the first step before the recursion to obtain speedups with up to 7 threads (see results section).
+
+\subsection{Solution to Linear Systems of Equations}
+
+Describe the solution to the linear system $Ax=b$ and backsubstitution. Describe Gaussian elimination with partial-pivoting.
+
+\section{Results}
+\subsection{Matrix-Vector Multiplication}
+
+\begin{center}
+\begin{tabular}{|r|r|r|r|r|r|r|}\hline
+{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
+1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+\end{tabular}
+\end{center}
+
+\subsection{Matrix-Matrix Multiplication}
+
+\begin{center}
+\begin{tabular}{|r|r|r|r|r|r|r|}\hline
+{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
+1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+\end{tabular}
+\end{center}
+
+\subsection{Gaussian Elimination and Solution}
+
+\begin{center}
+\begin{tabular}{|r|r|r|r|r|r|r|}\hline
+{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
+1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
+100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
+100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+\end{tabular}
+\end{center}
+
+\section{Conclusions}
+
+While we did blah, blah, blah and blah, blah, we recognize that blah is blah.
+For further development, we would need to implement blah blah blah.
+However, with regard to the library in its current state, we are satisfied with blah and would like to continue to blah.
+
 \newpage

### 2011-02-14 paper/paper.tex
Add all images and data tables

@@
 {\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
-1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & seq     &  22        & xxxxx       & xxxxx   & xxxxx     \\\hline
+1000    & 1       &  47        & 0.468       & 0.468   & xxxxx     \\\hline
+1000    & 2       &  40        & 0.550       & 0.275   & 0.702     \\\hline
+1000    & 4       &  34        & 0.647       & 0.162   & 0.631     \\\hline
+1000    & 8       &  35        & 0.629       & 0.079   & 0.708     \\\hline
+2000    & seq     &  83        & xxxxx       & xxxxx   & xxxxx     \\\hline
+2000    & 1       &  90        & 0.922       & 0.922   & xxxxx    \\\hline
+2000    & 2       &  35        & 2.371       & 1.186   & -0.222    \\\hline
+2000    & 4       &  14        & 5.929       & 1.482   & -0.126    \\\hline
+2000    & 8       &   9        & 9.222       & 1.153   & -0.029    \\\hline
+4000    & seq     &  338       & xxxxx       & xxxxx   & xxxxx     \\\hline
+4000    & 1       &  340       & 0.994       & 0.994   & xxxxx     \\\hline
+4000    & 2       &  177       & 1.910       & 0.955   & 0.041     \\\hline
+4000    & 4       &  91        & 3.714       & 0.929   & 0.024     \\\hline
+4000    & 8       &  34        & 9.941       & 1.243   & -0.029    \\\hline
 \end{tabular}
@@
 
+Sequential and Parallel Implementation
 \begin{center}
@@
 {\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
-1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & seq     &   32609        &   xxx         & xxx       & xxx        \\\hline
+1000    & 1       &   53546        & 0.609        & 0.609    & xxx        \\\hline
+1000    & 2       &   15812        & 2.062        & 1.031    & -0.409     \\\hline
+1000    & 4       &    8101        & 4.025        & 1.006    & -0.132     \\\hline
+1000    & 8       &    4703        & 6.934        & 0.991    & -0.064     \\\hline
+1000    & 8       &    4210        & 7.746        & 0.968    & -0.053     \\\hline
+500     & seq     &    4797        &   xxx         & xxx       & xxx        \\\hline
+500     & 1       &    2046        & 2.345         & 2.345     & xxx        \\\hline
+500     & 2       &    2632        & 1.823         & 0.911      & 1.573      \\\hline
+500     & 4       &    1411        & 3.400         & 0.850     & 0.586      \\\hline
+500     & 7       &     953        & 5.034         & 0.719     & 0.377      \\\hline
+500     & 8       &     789        & 6.080         & 0.760     & 0.298      \\\hline
+100     & seq     &      71        &   xxx         & xxx       & xxx        \\\hline
+100     & 1       &      62        & 1.145        & 1.145    & xxx        \\\hline
+100     & 2       &      42        & 1.690        & 0.845    & 0.355     \\\hline
+100     & 4       &      36        & 1.972        & 0.493    & 0.441     \\\hline
+100     & 7       &      35        & 2.029        & 0.290    & 0.492     \\\hline
+100     & 8       &      35        & 2.029        & 0.254    & 0.502     \\\hline
+\end{tabular}
+\end{center}
+
+Sequential Strassen and Parallel Strassen Implementation
+\begin{center}
+\begin{tabular}{|r|r|r|r|r|r|r|}\hline
+{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
+1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
+1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
+1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
+1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
+1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
+1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
+500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
+500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
+500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
+500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
+500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
+500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
+100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
+100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
+100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
+100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
+100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
+100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
 \end{tabular}
@@
 {\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
-1000    & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-1000    & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-1000    & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-1000    & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-500     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-500     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-500     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & seq     & 0000000        &   xxx         & xxx       & xxx        \\\hline
-100     & 1       & 0000000        & 0.0000        & 0.0000    & xxx        \\\hline
-100     & 2       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 4       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 8       & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
-100     & 16      & 0000000        & 0.0000        & 0.0000    & 0.0000     \\\hline
+1000    & seq     &  5374          & xxxxx        & xxxxx    & xxxxx     \\\hline
+1000    & 1       &  4412          & 1.218        & 1.218    & xxxxx     \\\hline
+1000    & 2       &  2393          & 2.246        & 1.123    & 0.085     \\\hline
+1000    & 4       &  1392          & 3.861        & 0.965    & 0.087     \\\hline
+1000    & 7       &  960           & 5.598        & 0.800    & 0.087     \\\hline
+1000    & 8       &  1851          & 2.903        & 0.363    & 0.337     \\\hline
+500     & seq     &  475           & xxxxx        & xxxxx    & xxxxx     \\\hline
+500     & 1       &  597           & 0.796        & 0.796    & xxxxx     \\\hline
+500     & 2       &  372           & 1.277        & 0.638    & 0.246     \\\hline
+500     & 4       &  249           & 1.908        & 0.477    & 0.223     \\\hline
+500     & 7       &  222           & 2.140        & 0.306    & 0.267     \\\hline
+500     & 8       &  324           & 1.466        & 0.183    & 0.477     \\\hline
+100     & seq     &  17            & xxxxx        & xxxxx    & xxxxx     \\\hline
+100     & 1       &  57            & 0.298        & 0.298    & xxxxx     \\\hline
+100     & 2       &  64            & 0.266        & 0.133    & 1.246     \\\hline
+100     & 4       &  66            & 0.258        & 0.064    & 1.211     \\\hline
+100     & 7       &  80            & 0.212        & 0.030    & 1.471     \\\hline
+100     & 8       &  101           & 0.168        & 0.021    & 1.882     \\\hline
 \end{tabular}

### 2011-02-15 paper/paper.tex
Second revision.

--- 2011-02-14 paper/paper.tex
@@
 
-The use of computers to manipulate and solve matrices has found many applications in computer science, finances, engineering, biology, and physics.
-Since many problems can be reduced to a system of linear equations, particularly in the solution of differential equations, numerical linear algebra has become increasingly useful in the study of such systems.
-In the High Performance Computing world, a large amount of effort has been devoted to developing toolsets that allow these problems to be solved in parallel.
+The use of computers to manipulate and solve matrices has found many applications in computer science, finances, engineering, biology, and physics \cite{TrefethenNLA}.
+Since many problems can be reduced to a system of linear equations, particularly in the solution of differential equations as seen in many examples in the Matrix Market maintained by NIST \cite{MatrixMarket}, numerical linear algebra has become increasingly useful in the study of such systems.
+In the High Performance Computing (HPC) world, a large amount of effort and resources have been devoted to developing toolsets that allow these problems to be solved in parallel \cite{SCALAPACK,LAPACK,PLASMA,PSBLAS}.
 
@@
 Early in the applications of computers to such problems, vector computers reigned supreme in numerical linear algebra due to their ability to apply operations to one-dimensional arrays of data.
-As scalar processors became cheaper and more readily available, a trend began of using large systems of scalar processors to perform parallel operations.
+As scalar processors became cheaper and more readily available, a trend began of using large networks of scalar processors to perform parallel operations.
+This trend continues today as we see some of the largest HPCs are based off of this model \cite{Jaguar}.
+
+While there are some leaders in the pack of toolsets providing an easily accessible parallel implementation of the standard BLAS \cite{SCALAPACK,PLAPACK}, no clear winner has emerged.
+The PMML does not attempt to challenge these toolsets, as such a challenger would be many years in development, but serves as an educational exercise in developing basic matrix manipulation methods in parallel.
 
@@
 
-Mention Parallel Java Library and need for Matrix Manipulation methods.
+The process of fitting more and more cores onto a single die is now leading us to consumer systems with multi-core chips.
+Since these systems share the symmetric multiprocessing model, these types of systems and larger SMP systems are the ideal target for the Parallel Matrix Manipulation Library (PMML).
+The PMML uses Parallel Java as a backbone to acheive the desired parallelism by dividing up computation in the agenda parallelism scheme.
+
+By dividing up the operations of multiplying slices of matrices among processors, a clear advantage is gained in terms of overall computation time.
+For the parallel version of the Strassen algorithm, the task of computing the block submatrices is divided among threads to achieve a reduced runtime.
+In the case of Gaussian elimination with partial-pivoting, the elementary row operations performed to reduce the matrix to row-echelon form are parallelized to achieve the desired speedup.
 
 \section{Methods}
+
+\subsection{Matrix-Matrix Multiplication}
+
+\subsubsection{Classic matrix multiplication}
+%Matrices of the form $A_{\mbox{\tiny m}\scriptscriptstyle\times\mbox{\tiny n}}$, where $m$ is the number of rows and $n$ is the number of columns.
+The Parallel Matrix Manipulation Library provides two methods for the multiplication of matrices of the form $A_{\mbox{\tiny m,n}}$, where $m$ is the number of rows and $n$ is the number of columns.
+The first of which is classic matrix multiplication of the form
+\begin{equation}
+\label{AB=C}
+A_{\mbox{\tiny m,n}}\cdot B_{\mbox{\tiny n,p}} = C_{\mbox{\tiny m,p}}
+\end{equation}
+where the product of an $m\times n$ matrix and a $n\times p$ matrix is an $m\times p$ matrix.
+
+In the PMML, the following code:
+\begin{verbatim}
+Matrix C = MatrixOp.mult(A,B);
+Matrix C = MatrixParallelOp.mult(A,B);
+\end{verbatim}
+returns the product $C=AB$, using the sequential and parallel algorithms respectively.
+Note that the operation is valid only if the number of columns of $A$ equals the number of rows of $B$, otherwise an Exception is thrown.
+
+The matrix multiplication is performed by multiplying each row of $A$ by each column of $B$ \cite{MatrixMultiplication}.
+Sequentially, one thread performs all the operations by looping on every row of $A$ and every column of $B$.
+The complexity is in $O(m\cdot n\cdot p)$ if $A$ and $B$ are rectangular matrices of size $m\times n$ and $n\times p$ respectively.
+This method is easily parallelizable over several threads; each thread computing a range of rows of $A$.
+The complexity is then $O( \frac{m}{K}\cdot n\cdot p )$, $K$ being the number of threads used.
+
+\subsubsection{Strassen matrix multiplication}
+The Parallel Matrix Manipulation Library also provides an implementation of the Strassen algorithm, an algorithm discovered by Volker Strassen that is $O( n^{2.807})$ \cite{StrassenAlgorithm}.
+
+Let $A$, $B$ be two real valued square matrices of size $2^n\times2^n$ for some $n$.
+If they are not this size, the missing rows and columns must be filled with zeros.
+In order to divide the computation, $A$, $B$ and $C$ are partitioned into equally sized blocks,
+\begin{equation*}
+\label{StrassenBlocks}
+\mathbf{A} = \begin{bmatrix} \mathbf{A}_{1,1} & \mathbf{A}_{1,2} \\ \mathbf{A}_{2,1} & \mathbf{A}_{2,2} \end{bmatrix}
+\mbox { , }
+\mathbf{B} = \begin{bmatrix} \mathbf{B}_{1,1} & \mathbf{B}_{1,2} \\ \mathbf{B}_{2,1} & \mathbf{B}_{2,2} \end{bmatrix}
+\mbox { , }
+\mathbf{C} = \begin{bmatrix} \mathbf{C}_{1,1} & \mathbf{C}_{1,2} \\ \mathbf{C}_{2,1} & \mathbf{C}_{2,2} \end{bmatrix}
+\end{equation*}
+with each $\mathbf{A}_{i,j}, \mathbf{B}_{i,j}, \mathbf{C}_{i,j}$ of size $2^{n-1} \times 2^{n-1}$. With this partitioning, the matrix $C$ can be expressed as the blocks
+\begin{eqnarray*}
+\mathbf{C}_{1,1} = \mathbf{A}_{1,1}\cdot \mathbf{B}_{1,1} + \mathbf{A}_{1,2}\cdot \mathbf{B}_{2,1},\\
+\mathbf{C}_{1,2} = \mathbf{A}_{1,1}\cdot \mathbf{B}_{1,2} + \mathbf{A}_{1,2}\cdot \mathbf{B}_{2,2},\\
+\mathbf{C}_{2,1} = \mathbf{A}_{2,1}\cdot \mathbf{B}_{1,1} + \mathbf{A}_{2,2}\cdot \mathbf{B}_{2,1},\\
+\mathbf{C}_{2,2} = \mathbf{A}_{2,1}\cdot \mathbf{B}_{1,2} + \mathbf{A}_{2,2}\cdot \mathbf{B}_{2,2}.
+\end{eqnarray*}
+In the case of standard matrix multiplication, the partitioning of $C$ in this way yields a total of 8 matrix multiplications. By introducing the following matrices
+\begin{align*}
+&\mathbf{M}_{1} := (\mathbf{A}_{1,1} + \mathbf{A}_{2,2})\cdot (\mathbf{B}_{1,1} + \mathbf{B}_{2,2}),\\
+&\mathbf{M}_{2} := (\mathbf{A}_{2,1} + \mathbf{A}_{2,2})\cdot \mathbf{B}_{1,1},\\
+&\mathbf{M}_{3} := \mathbf{A}_{1,1}\cdot (\mathbf{B}_{1,2} - \mathbf{B}_{2,2}),\\
+&\mathbf{M}_{4} := \mathbf{A}_{2,2}\cdot (\mathbf{B}_{2,1} - \mathbf{B}_{1,1}),\\
+&\mathbf{M}_{5} := (\mathbf{A}_{1,1} + \mathbf{A}_{1,2})\cdot \mathbf{B}_{2,2},\\
+&\mathbf{M}_{6} := (\mathbf{A}_{2,1} - \mathbf{A}_{1,1})\cdot (\mathbf{B}_{1,1} + \mathbf{B}_{1,2}),\\
+&\mathbf{M}_{7} := (\mathbf{A}_{1,2} - \mathbf{A}_{2,2})\cdot (\mathbf{B}_{2,1} + \mathbf{B}_{2,2}).
+\end{align*}
+and then computing 
+\begin{align*}
+&\mathbf{C}_{1,1} = \mathbf{M}_{1} + \mathbf{M}_{4} - \mathbf{M}_{5} + \mathbf{M}_{7}\\
+&\mathbf{C}_{1,2} = \mathbf{M}_{3} + \mathbf{M}_{5}\\
+&\mathbf{C}_{2,1} = \mathbf{M}_{2} + \mathbf{M}_{4}\\
+&\mathbf{C}_{2,2} = \mathbf{M}_{1} - \mathbf{M}_{2} + \mathbf{M}_{3} + \mathbf{M}_{6}
+\end{align*}
+we see that the total number of matrix partition multiplications drops from $8$ to $7$.
+Since these matrix multiplications are more costly than matrix additions and subtractions, efficiency is gained in the computation of $C$.
+
+The idea consists in recursively dividing both matrices into 4 smaller square matrices, until those submatrices degenerate into numbers, that is have a size of 1, and replace the last multiplication by the sum of 18 submatrices.
+However, by padding with zeroes the matrices to the closest power of 2, we can multiply matrices of any size.
+Due to this recursive nature, this algorithm is not easily parallelizable but we managed to parallelize the first step before the recursion to obtain speedups with up to 7 threads (see results section).
+
+In the PMML, the following code:
+
+{\tt Matrix C = MatrixOp.strassenMult(A,B);\\
+Matrix C = MatrixParallelStr.mult(A,B);}
+
+returns the product $C=AB$, using the sequential and parallel versions respectively of the Strassen algorithm.
+
 \subsection{Matrix-Vector Multiplication}
@@
 
-\subsection{Matrix-Matrix Multiplication}
-
-\subsubsection{Classic matrix multiplication}
-The following code:
-\begin{verbatim}
-Matrix C = MatrixOp.mult(A,B)
-Matrix C = MatrixParallelOp.mult(A,B)
-\end{verbatim}
-returns the product $C=AB$, using respectively the sequential and parallel algorithms.
-
-Note that the operation is valid only if the number of columns of $A$ equals the number of rows of $B$, otherwise an Exception is thrown.
-
-The matrix multiplication is performed by multiplying each row of $A$ by each column of $B$ \cite{MatrixMultiplication}.
-Sequentially, one thread performs all the operations by looping on every rows of $A$ and every columns of $B$.
-The complexity is in $O(m\cdot p\cdot n)$ if $A$ and $B$ are rectangular matrices of size $m\cdot p$ and $p\cdot n$ respectively.
-This method is easily parallelizable over several threads; each thread computing a range of rows of $A$.
-The complexity is then $O( \frac{m}{K}\cdot p\cdot n )$, $K$ being the number of threads used.
-
-\subsubsection{Strassen matrix multiplication}
-The following code:
-\begin{verbatim}
-Matrix C = MatrixOp.strassenMult(A,B)
-Matrix C = MatrixParallelStr.mult(A,B)
-\end{verbatim}
-returns the product $C=AB$, using respectively the sequential and parallel versions of the Strassen algorithms.
-
-There is a way of multiplying two $2\times2$ matrices using only 7 multiplications instead of 8 with the classic algorithm.
-The Strassen algorithm takes advantage of that to multiply two matrices with a complexity of $O(n^2.807)$, providing that those matrices have the same size and are square matrices with side of size a power of 2 \cite{StrassenAlgorithm}.
-The idea consists in recursively dividing both matrices into 4 smaller square matrices, until those submatrices degenerate into numbers, that is have a size of 1, and replace the last multiplication by the sum of 18 submatrices.
-
-However, by padding with zeroes the matrices to the closest power of 2, we can multiply matrices of any size.
-
-Due to this recursive nature, this algorithm is not easily parallelizable but we managed to parallelize the first step before the recursion to obtain speedups with up to 7 threads (see results section).
-
 \subsection{Solution to Linear Systems of Equations}
 
-Describe the solution to the linear system $Ax=b$ and backsubstitution. Describe Gaussian elimination with partial-pivoting.
+To solve a linear system of equations of the form $Ax=b$, where $A$ is an $m\times n$ matrix, $b$ is an $m\times 1$ vector, and $x$ is the $m\times 1$ solution vector, first the matrix must be inverted.
+For most matrices, this presents a problem, as the inverse of a matrix is generally not easy to compute and may not even exist.
+Gaussian elimination is the process of performing a series of elementary row operations to the matrix $A$ to reduce it to a form, known as row-echelon form, that can be easily inverted.
 

### 2011-02-15 paper/paper.tex
Merge

@@
 
+\begin{verbatim}
+MatrixInt A = ...
+VectorInt B = ...
+Matrix C = MatrixOp.mult(A,B)
+Matrix C = MatrixParallelOp.mult(A,B)
+
+\end{verbatim}
+
 \subsection{Matrix-Matrix Multiplication}
@@
 \section{Results}
+
 \subsection{Matrix-Vector Multiplication}
@@
 \end{tabular}
+\end{center}
+
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-vector/running-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-vector/speedup-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-vector/esdf-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-vector/eff-vs-proc.png}
 \end{center}
@@
 
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel/run-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel/speedup-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel/edsf-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel/eff-vs-proc.png}
+\end{center}
+
 Sequential Strassen and Parallel Strassen Implementation
@@
 \end{tabular}
+\end{center}
+
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel-stras/running-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel-stras/speedup-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel-stras/edsf-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/parallel-stras/eff-vs-proc.png}
 \end{center}
@@
 
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-inversion/running-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-inversion/speedup-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-inversion/edsf-vs-proc.png}
+\end{center}
+\begin{center}
+\includegraphics[scale=0.5]{images/matrix-inversion/eff-vs-proc.png}
+\end{center}
+
 \section{Conclusions}

### 2011-02-15 paper/paper.tex
Added proper description to Gaussian elimination.

@@
 
-Describe the problem of matrix-vector multiplication and introduce the terms.
+In the case of matrix-vector multiplication $A\cdot v$, the vector $\vec{v}$ is treated as a matrix of size $n\times 1$.
+As $\vec{v}$ is not of requisite size for the Strassen algorithm to be applied, only standard matrix-vector multiplication is used.
+Thus the process of computing this product is parallelized by dividing up the computation of dot products with the rows of $A$ and $\vec{v}$ among the threads.
 
@@
 
-To solve a linear system of equations of the form $Ax=b$, where $A$ is an $m\times n$ matrix, $b$ is an $m\times 1$ vector, and $x$ is the $m\times 1$ solution vector, first the matrix must be inverted.
+To solve a linear system of equations of the form 
+\begin{equation}
+$Ax=b$, 
+\end{equation}
+where $A$ is an $m\times n$ matrix, $b$ is an $m\times 1$ vector, and $x$ is the $m\times 1$ solution vector, first the matrix must be inverted.
 For most matrices, this presents a problem, as the inverse of a matrix is generally not easy to compute and may not even exist.
 Gaussian elimination is the process of performing a series of elementary row operations to the matrix $A$ to reduce it to a form, known as row-echelon form, that can be easily inverted.
+The matrix
+\begin{equation} 
+\left[ \begin{array}{ccc|c}
+a_{1,1} & a_{1,2} & a_{1,3} & b_1 \\
+0 & a_{2,2} & a_{2,2} & b_2 \\
+0 & 0 & a_{3,3} & b_3
+\end{array} \right],
+\end{equation}
+where $a_{i,j}$ are the elements of $A$, and $b_i$ are the elements of $b$, is an example of the system $Ax=b$ in row-echelon form.
+Matrices in row-echelon form have the nice characteristic that all of the leading coefficients of each row are in a column whose number is less than the column of the leading coefficients in the rows below it and greater than those in the rows above it.
+In the process of Gaussian elimination with partial-pivoting, this is achieved by moving the row with the greatest leading coefficient above those with lesser coefficients, and then reducing the lower rows.
+In the case of square matrices whose rows are linearly independent, this reduction results in an upper-triangular form.
+Once the matrix is in this form, the inversion process is equivalent to back substituting the solution from the bottom of the matrix all the way to the top.
 

### 2011-02-15 paper/paper.tex
Added more to Gaussian elimination and some minor changes.

@@
 Once the matrix is in this form, the inversion process is equivalent to back substituting the solution from the bottom of the matrix all the way to the top.
+Parallelization of this method consists of dividing up the task of reducing the rows of the matrix below the pivot point among threads.
 
 \section{Results}
+
+To test each of these methods, in the case of providing a matrix, a random matrix of size $n^2$ was generated by:
+\begin{verbatim}
+MatrixDouble A = MatrixDouble.random(1000, 1000);
+\end{verbatim}
+and in the case of a vector, a random vector of size $n$ was generated by:
+\begin{verbatim}
+VectorDouble b = VectorDouble.random(1000);
+\end{verbatim}
+Over 4 runs, the time required to perform the operation was averaged.
 
@@
 
-Sequential and Parallel Implementation
+\subsubsection{Sequential and Parallel Implementation}
 \begin{center}
@@
 
-Sequential Strassen and Parallel Strassen Implementation
+\subsubsection{Sequential Strassen and Parallel Strassen Implementation}
 \begin{center}

### 2011-02-15 src/MatrixDouble.java
Small beatification of the code and adding/fixing documentation

@@ -18,20 +18,19 @@ public class MatrixDouble {
 
     data = new double[rows][cols];
   }
-  
+
   /** return the identity matrix
    * @param size the size of a side of a matrix (identity->square matrix)
    * @return the identity matrix of size (sixe*sixe)
    */
-  public static MatrixDouble getIdentity(int size)
-  {
-	  double[][] data = new double[size][size];
-	  // should be initialized at 0 by default
-	  for(int i=0; i<size; i++)
-		  data[i][i] = 1.0f;
-	  
-	  MatrixDouble m = new MatrixDouble(data);
-	  return m;
+  public static MatrixDouble getIdentity(int size) {
+    double[][] data = new double[size][size];
+    // should be initialized at 0 by default
+    for(int i=0; i<size; i++)
+      data[i][i] = 1.0f;
+
+    MatrixDouble m = new MatrixDouble(data);
+    return m;
   }
 
   /**
@@ -142,9 +141,9 @@ public class MatrixDouble {
    */
   public boolean correctDim( MatrixDouble b)
   {
-	  return this.cols() == b.rows();
+    return this.cols() == b.rows();
   }
-  
+
   /** determine whether the dimension of the 2 matrices allows us to perform a matrix multiplication 
    * @param this the left matrix
    * @param b the right matrix
@@ -152,9 +151,9 @@ public class MatrixDouble {
    */
   public boolean correctDim( MatrixInt b)
   {
-	  return this.cols() == b.rows();
+    return this.cols() == b.rows();
   }
-  
+
   /**
    * Member function to display the current matrix on the command line.
    */

### 2011-02-15 src/MatrixInt.java

@@ -35,8 +35,6 @@ public class MatrixInt {
     data = new int[rows][cols];
     this.copyData( cdata );
   }
-  
-  
 
   /**
    * Matrix copy constructor
@@ -44,23 +42,22 @@ public class MatrixInt {
   private MatrixInt( final MatrixInt copy ) { 
     this( copy.data );
   }
-       
-  
+
   /** return the identity matrix
    * @param size the size of a side of a matrix (identity->square matrix)
    * @return the identity matrix of size (sixe*sixe)
    */
   public static MatrixInt getIdentity(int size)
   {
-	  int[][] data = new int[size][size];
-	  // should be initialized at 0 by default
-	  for(int i=0; i<size; i++)
-		  data[i][i] = 1; 
-	  
-	  MatrixInt m = new MatrixInt(data);
-	  return m;
+    int[][] data = new int[size][size];
+    // should be initialized at 0 by default
+    for(int i=0; i<size; i++)
+      data[i][i] = 1; 
+
+    MatrixInt m = new MatrixInt(data);
+    return m;
   }
-  
+
   /** Get the number of rows in the matrix */
   public int rows() { return rows; }
 
@@ -151,7 +148,7 @@ public class MatrixInt {
    */
   public boolean correctDim( MatrixDouble b)
   {
-	  return this.cols() == b.rows();
+    return this.cols() == b.rows();
   }
   
   /** determine whether the dimension of the 2 matrices allows us to perform a matrix multiplication 
@@ -161,7 +158,7 @@ public class MatrixInt {
    */
   public boolean correctDim( MatrixInt b)
   {
-	  return this.cols() == b.rows();
+    return this.cols() == b.rows();
   }
 
   /**
@@ -187,5 +184,5 @@ public class MatrixInt {
     MatrixInt r = MatrixInt.randomSmall( 20, 20 );
     r.display();
   }
-  
+
 }

### 2011-02-15 src/MatrixInverse.java

@@ -1,7 +1,7 @@
 import edu.rit.pj.Comm;
 
 /**
- * A program that applies Gaussian eliminati n t  compute
+ * A program that applies Gaussian elimination to compute
  * the inverse that solves the system Ax=b
  */
 public class MatrixInverse{
@@ -75,8 +75,10 @@ public class MatrixInverse{
         return x;
     }
 
-    /** 
-     * apply gaussian elimination to the matrix a
+    /**
+     * Apply gaussian elimination to the matrix a
+     *
+     * @param A - Matrix to eliminate 
      */
     public static MatrixDouble eliminate(MatrixDouble A) throws Exception {
         // Determine the number of rows of A
@@ -122,8 +124,8 @@ public class MatrixInverse{
         return A;
     }
 
-    /** 
-     * main method for testing the inversion
+    /**
+     * Main method for testing the inversion
      */
     public static void main(String args[]) throws Exception {
 
@@ -170,4 +172,4 @@ public class MatrixInverse{
                 (System.currentTimeMillis()-start)/loop + " msec" );
     }
 
-}//MatrixInverse.java
+} //MatrixInverse.java

### 2011-02-15 src/MatrixInverseSmp.java

@@ -4,7 +4,7 @@ import edu.rit.pj.ParallelTeam;
 import edu.rit.pj.IntegerForLoop;
 
 /**
- * A program that applies Gaussian eliminati n t  compute
+ * A smp program that applies Gaussian elimination to compute
  * the inverse that solves the system Ax=b
  */
 public class MatrixInverseSmp{
@@ -14,8 +14,6 @@ public class MatrixInverseSmp{
 
     // prevent construction
     private MatrixInverseSmp() { }
-
-
 
     /**
      * Solves the linear system A.x=b
@@ -185,7 +183,7 @@ public class MatrixInverseSmp{
             System.err.println(re.getMessage());
         }
         */
-        
+
         int loop = 10;
         long start = System.currentTimeMillis();
         for(int i = 0; i < loop; i++) {

### 2011-02-15 src/MatrixOp.java

@@ -70,7 +70,7 @@ public class MatrixOp {
   {
 	  if ( ! a.equalDim( b ) ) 
 	  {
-	      throw new RuntimeException("Multiplication of different sized matricies");
+	      throw new RuntimeException("Addition of different sized matricies");
 	  }
 	  int[][] data = new int[a.rows()][a.cols()];
 	  for(int i=0; i<a.rows(); i++ )
@@ -94,7 +94,7 @@ public class MatrixOp {
   {
 	  if ( ! a.equalDim( b ) ) 
 	  {
-	      throw new RuntimeException("Multiplication of different sized matricies");
+	      throw new RuntimeException("Subtraction of different sized matricies");
 	  }
 	  int[][] data = new int[a.rows()][a.cols()];
 	  for(int i=0; i<a.rows(); i++ )
@@ -280,7 +280,6 @@ public class MatrixOp {
    */
   public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
   {
-	//System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
     int[][] data = new int[nbrows][nbcols];
     for(int i = 0; i < a.rows(); i++ ) {
       for(int j = 0; j < a.cols(); j++ ) {
@@ -299,68 +298,60 @@ public class MatrixOp {
 
   public static MatrixInt strassenMult( MatrixInt a, MatrixInt b )
   {
-	  if ( ! a.correctDim( b ) ) {
-	      throw new RuntimeException("Multiplication of different sized matricies");
-	  }
-	  
+    if ( ! a.correctDim( b ) ) {
+        throw new RuntimeException("Multiplication of different sized matricies");
+    }
+
     // check if the size is a power of 2 :
-	int finalrows = a.rows();
-	int finalcol = b.cols();
-	  
+    int finalrows = a.rows();
+    int finalcol = b.cols();
+
     int rows = Math.max(a.rows(),b.rows());
     int cols = Math.max(a.cols(),b.cols());
     int resizing = Math.max(rows,cols);
     // we want a and b to be 2 matrices of size the closest power of 2 of 'resizing'
-	
+
     MatrixInt result;
 
     if ( a.equalDim(b)  && isPow2(rows) ) 
     {
-        //System.out.println("no need to pad");
-    	result =  strassenMultRun(a, b);
-
[…]

### 2011-02-15 src/MatrixParallelStr.java

@@ -14,16 +14,14 @@ public class MatrixParallelStr {
   // Create a static team to avoid allocation hit
   static ParallelTeam team = new ParallelTeam();
 
-  
+
   /** pads a matrix with 0 until it reaches the new size nbrows, nbcols
    * 
    * @param a
    * @param nbrows
    * @param nbcols
    */
-  public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) 
-  {
-	//System.out.println("resizing from "+a.rows()+","+a.cols()+" to " +nbrows+ ","+nbcols);
+  public static MatrixInt pad( MatrixInt a, int nbrows, int nbcols ) {
     int[][] data = new int[nbrows][nbcols];
     for(int i = 0; i < a.rows(); i++ ) {
       for(int j = 0; j < a.cols(); j++ ) {
@@ -38,54 +36,43 @@ public class MatrixParallelStr {
   {
     return ((number ^ (number-1)) == (number+(number-1))) && (number != 0);
   }
-  
-  public static MatrixInt mult( MatrixInt a, MatrixInt b ) throws Exception 
-  {
-	  
-	  if ( ! a.correctDim( b ) ) {
-	      throw new RuntimeException("Multiplication of different sized matricies");
-	  }
-	  
+
+  public static MatrixInt mult( MatrixInt a, MatrixInt b ) throws Exception {
+
+    if ( ! a.correctDim( b ) ) {
+        throw new RuntimeException("Multiplication of different sized matricies");
+    }
+
     // check if the size is a power of 2 :
-	int finalrows = a.rows();
-	int finalcol = b.cols();
-	  
+  int finalrows = a.rows();
+  int finalcol = b.cols();
+
     int rows = Math.max(a.rows(),b.rows());
     int cols = Math.max(a.cols(),b.cols());
     int resizing = Math.max(rows,cols);
     // we want a and b to be 2 matrices of size the closest power of 2 of 'resizing'
-	
+
     MatrixInt result;
 
-    if ( a.equalDim(b)  && isPow2(rows) ) 
-    {
-        //System.out.println("no need to pad");
-    	result =  multRunSmp( a, b );
-
-    }
-    else
-    {
-    	if( ! isPow2(resizing) )
+    if ( a.equalDim(b)  && isPow2(rows) ) {
+      result =  multRunSmp( a, b );
+    } else {
+      if( ! isPow2(resizing) )
[…]

### 2011-02-15 src/Tester.java

@@ -2,7 +2,7 @@ import edu.rit.pj.Comm;
 
 public class Tester {
 
-	  static int NUM_RUNS = 4;
+    static int NUM_RUNS = 4;
 
     private static void runTest( int N ) throws Exception {
 
@@ -11,9 +11,9 @@ public class Tester {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
           MatrixInt C = MatrixOp.mult( A, B );
-				}
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
@@ -26,9 +26,9 @@ public class Tester {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
-        	MatrixInt C = MatrixOp.strassenMult( A, B );
-				}
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixOp.strassenMult( A, B );
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Strassen ("+N+") = "+(end-start)/NUM_RUNS);
@@ -41,9 +41,9 @@ public class Tester {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
-        	MatrixInt C = MatrixParallelOp.mult( A, B );
-				}
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixParallelOp.mult( A, B );
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Parallel ("+N+") = "+(end-start)/NUM_RUNS);
@@ -56,9 +56,9 @@ public class Tester {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
-        	MatrixInt C = MatrixParallelStr.mult( A, B );
-				}
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixParallelStr.mult( A, B );
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Parallel Strassen ("+N+") = "+(end-start)/NUM_RUNS);
@@ -70,21 +70,21 @@ public class Tester {
[…]

### 2011-02-15 src/TesterInverse.java

@@ -2,7 +2,7 @@ import edu.rit.pj.Comm;
 
 public class TesterInverse {
 
-	  static int NUM_RUNS = 4;
+    static int NUM_RUNS = 4;
 
     private static void runTest( int N ) throws Exception {
 
@@ -12,9 +12,9 @@ public class TesterInverse {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
           VectorDouble C = MatrixInverse.solve( A, B );
-				}
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
@@ -27,9 +27,9 @@ public class TesterInverse {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
-        	VectorDouble C = MatrixInverseSmp.solve( A, B );
-				}
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
+          VectorDouble C = MatrixInverseSmp.solve( A, B );
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Parallel ("+N+") = "+(end-start)/NUM_RUNS);

### 2011-02-15 src/TesterVector.java

@@ -2,7 +2,7 @@ import edu.rit.pj.Comm;
 
 public class TesterVector {
 
-	  static int NUM_RUNS = 4;
+    static int NUM_RUNS = 4;
 
     private static void runTest( int N ) throws Exception {
 
@@ -11,9 +11,9 @@ public class TesterVector {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
           MatrixInt C = MatrixOp.mult( A, B );
-				}
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Classic = ("+N+") = " +(end-start)/NUM_RUNS);
@@ -26,9 +26,9 @@ public class TesterVector {
 
         long start = System.currentTimeMillis();
 
-				for ( int i = 0; i < NUM_RUNS; i++ ) {
-        	MatrixInt C = MatrixParallelOp.mult( A, B );
-				}
+        for ( int i = 0; i < NUM_RUNS; i++ ) {
+          MatrixInt C = MatrixParallelOp.mult( A, B );
+        }
 
         long end = System.currentTimeMillis();
         System.out.println( "Parallel ("+N+") = "+(end-start)/NUM_RUNS);

### 2011-02-15 src/VectorDouble.java

@@ -5,6 +5,9 @@ public class VectorDouble {
     private int size;
     public double[] data;
 
+    /**
+     * Default constructor 
+     */
     public VectorDouble( int size ) {
         this.size = size;
         this.data = new double[size];
@@ -22,8 +25,8 @@ public class VectorDouble {
         this.copyData( cdata );
     }
 
-    /** 
-     * Copy the data from a given array into 
+    /**
+     * Copy the data from a given array into
      * the vector own allocated data.
      *
      * @param cdata The array to copy from
@@ -49,10 +52,16 @@ public class VectorDouble {
         return rand;
     }
 
+    /**
+     * Obtain the lenght of the vector
+     */
     public int length() {
         return  size;
     }
 
+    /**
+     * Zero out the current vector
+     */
     public void zero() {
         for ( int i = 0; i < size; i++ ) {
             this.data[i] = 0;

### 2011-02-15 src/VectorInt.java

@@ -10,10 +10,40 @@ public class VectorInt {
         this.data = new int[size];
     }
 
+    /**
+     * Data copy constructor
+     *
+     * @param cdata The data to copy
+     */
+    public VectorInt( int[] cdata ) {
+        this.size = cdata.length;
+
+        data = new int[size];
+        this.copyData( cdata );
+    }
+
+    /**
+     * Copy the data from a given array into
+     * the vector own allocated data.
+     *
+     * @param cdata The array to copy from
+     */
+    private void copyData( int[] cdata ) {
+        for ( int r = 0; r < size ; r++ ) {
+                this.data[r] = cdata[r];
+        }
+    }
+
+    /**
+     * Obtain the lenght of the vector
+     */
     public int length() {
         return  size;
     }
 
+    /**
+     * Zero out the current vector
+     */
     public void zero() {
         for ( int i = 0; i < size; i++ ) {
             this.data[i] = 0;
@@ -34,6 +64,5 @@ public class VectorInt {
         }
         return rand;
     }
-
 
 }

### 2011-02-15 paper/paper.tex
Add some explination's to the results

@@
 
-A Parallel Matrix Manipulation Library (PMML) is presented for use on sequential and symmetric multiprocessing (SMP) systems.
-PMML is written using the Parallel Java library developed by Professor Alan Kaminsky at the Rochester Institute of Technology.
+A Parallel Matrix Manipulation Library (PMML) is presented for use on
+sequential and symmetric multiprocessing (SMP) systems.
+PMML is written using the Parallel Java (PJ) library developed by Professor Alan Kaminsky at the Rochester Institute of Technology.
 PMML provides methods for the following operations: matrix-vector multiplication, matrix-matrix multiplication, and matrix inversion and solution to the linear system $A x=b$.
@@
 
-\begin{verbatim}
-MatrixInt A = ...
-VectorInt B = ...
-Matrix C = MatrixOp.mult(A,B)
-Matrix C = MatrixParallelOp.mult(A,B)
-
-\end{verbatim}
-
 \subsection{Matrix-Matrix Multiplication}
@@
 \begin{verbatim}
+MatrixInt A = ...
+MatrixInt B = ...
 Matrix C = MatrixOp.mult(A,B);
@@
 
+In the PMML, the following code:
+
+{\tt MatrixInt A = ...\\
+VectorInt v = ... \\
+Matrix C = MatrixOp.mult(A,v) \\
+Matrix C = MatrixParallelOp.mult(A,v) }
+
+returns the product $C=A\cdot v$, using the sequential and parallel versions of the
+matrix vector multiplication algorithm. Note the use of Java's polymorphism,
+allowing a user to multiply a vector \& matrix, as
+well as a matrix \& matrix using the same command.
+
 \subsection{Solution to Linear Systems of Equations}
@@
 \begin{equation}
-$Ax=b$, 
+Ax=b, 
 \end{equation}
@@
 To test each of these methods, in the case of providing a matrix, a random matrix of size $n^2$ was generated by:
-\begin{verbatim}
-MatrixDouble A = MatrixDouble.random(1000, 1000);
-\end{verbatim}
+
+{\tt MatrixDouble A = MatrixDouble.random(1000, 1000); }
+
 and in the case of a vector, a random vector of size $n$ was generated by:
-\begin{verbatim}
-VectorDouble b = VectorDouble.random(1000);
-\end{verbatim}
+
+{\tt VectorDouble b = VectorDouble.random(1000);}
+
 Over 4 runs, the time required to perform the operation was averaged.
+This technique was chosen in an attempt to force the JVM's Just in Time
+Compiler to activate and produce machine code for the methods
+being tested. This gives an average of the initial non-JIT timings, mixed
+with the eventual JIT-enabled timings.
 
@@
 \subsubsection{Sequential and Parallel Implementation}
+
 \begin{center}
@@
 \end{center}
-
 \begin{center}
@@
 \end{center}
+The results report strong parallel speedup especially as our matrix 
+size increases. The speedup for the $1000 X 1000$ matrix multiplication reports
+near linear speedup, and a efficiency of $1.0$ exactly. The particularly poor performance
+for small matrices can potentially be traced back to the slight over head
+that comes with starting and executing a Parallel Java team. It is possible that the time taken during
+the initial dispatch of the threads in the parallel implementation is greater
+than any possible speed up achieved by using them for small matrices.
 
-\subsubsection{Sequential Strassen and Parallel Strassen Implementation}
+\subsubsection{Sequential \& Parallel Strassen Implementation}
+
 \begin{center}
@@
 \end{center}
+Overall the results for the sequential Strassen and Parallel-Strassen
+algorithms reported significantly faster run-times when compared to the
+sequential and parallel regular algorithms. However the speedup
+compared to the other matrix multiplication results is not quite as
+impressive. Their are many factors which contribute to this result.
+First of all, for matrices smaller than size of 128 by 128 both of the
+Strassen implementations default back to the regular sequential matrix
+multiplication. During testing it was discovered that the regular sequential
+algorithm is faster for this special case. This is why the results show no
+speedup for 100 by 100 matrix. It should also be noted that in our parallel
+Strassen implementation the maximum number of steps which can parallelized is
+seven. As shown in the results above, when moving to eight parallel threads
+very little speedup is achieved. The Strassen algorithms are also specially
+designed to operate on matrices which are of a size which is a power of two.
+In order to take advantage of this for ever matrix, if the matrix is not a
+power of two the matrix is padded with zero's and then unpadded after the
+algorithm is complete. Since none of the matrices tested were of a size which
+was a power of two, the results shown are examples of the worst possible
+running time for the sequential and parallel Strassen algorithms.
 

### 2011-02-15 paper/paper.tex
Add some more explinations of the results

@@
 
+As shown above, the parallel matrix-vector multiplication's report super
+linear speedup's for large matrices and accompanying vectors. It should be
+noted that larger N's were used in testing these algorithms vs the results
+that follow. This was merely because this algorithm is far less memory and
+computationally intensive. The run times with a smaller N were so small that
+it was hard to discern if a speedup was occurring in some cases. It is possible,
+as seen in the other results through out this paper, that the poor performance
+for the smaller 1000 by 1000 matrix multiplied by a vector of length 1000 is
+caused by the overhead associated with Parallel Java. The time it takes to
+actually compute the result is shorter than the time it takes to initialize
+and dispatch the threads from Parallel Java.
+
 \subsection{Matrix-Matrix Multiplication}
@@
 The results report strong parallel speedup especially as our matrix 
-size increases. The speedup for the $1000 X 1000$ matrix multiplication reports
-near linear speedup, and a efficiency of $1.0$ exactly. The particularly poor performance
+size increases. The speedup for the 1000 by 1000 matrix multiplication reports
+near linear speedup, and a efficiency of 1.0 exactly. The particularly poor performance
 for small matrices can potentially be traced back to the slight over head
@@
 algorithm is faster for this special case. This is why the results show no
-speedup for 100 by 100 matrix. It should also be noted that in our parallel
+speedup for 100 by 100 matrix, and the first data point for the  100 by 100
+matrix has such a extremely high efficiency. It should also be noted that in our parallel
 Strassen implementation the maximum number of steps which can parallelized is
@@
 
+These results are quite interesting in many ways. As observed in all of the
+previous results, we see a increasing level of speedup as our problem size
+increases. With the largest N performing the best, with a linear
+speed up initially and slowly dropping in performance, until a drastic
+crash in performance can be observed with running eight threads. This could
+possibly be the final thread phenomenan discussed in class, where the
+performance of the algorithm is reduced when the number of threads is equal
+to the number of processors available. It would be interesting to continue
+testing larger N's and seeing if the effect is more drastic the larger the N
+as that appears to be what is happening with the relatively small data set.
 \section{Conclusions}

### 2011-02-15 paper/paper.tex
Add spec's of the machine the timing numbers were recorded on.
Add small section to conclusion

@@
 
+All timing numbers were taken by running jobs on the RIT CS department's
+paradise SMP machine. The machines stat's at the time of writing: 
+Four Sun UltraSPARC-IV dual-core CPUs, eight processors, 1.35 GHz clock, 16 GB
+main memory.
 \subsection{Matrix-Vector Multiplication}
@@
 
+We were able to glean interesting results from the data points we retreived.
+However the size of the matrices we could create was limited by JVM's maximum heap size.
+It may be worth wile to explore running the implementented
+algorithms on different architectures with possibly larger JVM heap upper
+bounds. Since in general we saw increased peformance by increasing the size of
+N, the point at which this is no longer true for our implementations would be a
+helpful thing to know.
+
 \newpage

### 2011-02-15 paper/paper.tex
Was editing the file and saw changes had been made, committimg my changes to merge

@@
 speedup for 100 by 100 matrix, and the first data point for the  100 by 100
-matrix has such a extremely high efficiency. It should also be noted that in our parallel
-Strassen implementation the maximum number of steps which can parallelized is
-seven. As shown in the results above, when moving to eight parallel threads
+matrix has such a extremely high efficiency. It should also be noted that in 
+our parallel Strassen implementation the maximum number of steps which can 
+parallelized is seven.
+
+As shown in the results above, when moving to eight parallel threads
 very little speedup is achieved. The Strassen algorithms are also specially
@@
 as that appears to be what is happening with the relatively small data set.
+
+One of the most difficult things to parallelize is the solution of a linear 
+system of equations. Usually, the properties of the matrix being used must 
+be exploited to receive any kind of speedup or efficiency gains. Gaussian 
+elimination in particular is not very parallelizable, especially with 
+partial-pivoting. The actual act of pivoting the rows of matrices introduces 
+a sequential dependency to the process. Still, the parallelization of the 
+elementary row operations proved to be beneficial to the overal performance 
+of the method.
+
 \section{Conclusions}
